CSE530 Project Final Report\\ Bandwidth-Aware Memory Hierarchy Design with Hybrid Memory Technologies

### 2010-11-18 paper.tex
combined version

\documentclass[10pt,letterpaper,singlecolumn]{article}
\usepackage{setspace}
\doublespacing
\usepackage{latex8}
\usepackage{times}
\usepackage{cite}
\usepackage{subfigure}
\usepackage{psfig}
\usepackage[pdftex]{graphicx}
\usepackage{epsfig}
\usepackage{pstricks}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{graphics}
\usepackage{comment}
\usepackage[small,compact]{titlesec}
\usepackage[small,it]{caption}
\usepackage{wrapfig}
\usepackage{multirow}

@@

\begin{figure*}[htbp]
% The "!" means to maintain the aspect ratio.
\centering
\includegraphics[width=6in]{figures/RAM_energy}
\caption{Dynamic energy consumption with the provided bandwidths of different
  memory technologies. (a) Dynamic Energy with 20\% write.  (b) Dynamic Energy with 50\% write. }
\label{fig:memory-energy}
\end{figure*}

\begin{figure*}[htbp]
% The "!" means to maintain the aspect ratio.
\centering
\includegraphics[width=6in]{figures/RAM_latency}
\caption{Latency of different memory technologies. (a) Read Latency. (b) Write Latency.}
\label{fig:latency}
\end{figure*}

First of all, we estimate the read and write bandwidths that can be provided by
different memory technologies. Figure~\ref{fig:memory-bw} shows the results,
with both x- and y-values in \emph{log} scale. The figure illustrates both the
provided read and write bandwidth as a function of memory capacity. Each of the
memory technologies actually provide nearly the same read bandwidths, as is
shown in figure~\ref{fig:memory-bw}(a). On the other hand, a straight forward
observation from figure~\ref{fig:memory-bw}(b) is that the write bandwidth
varies among different memory technologies. The shape of the SRAM write
bandwidth curve is very similar to the read bandwidth curve. The write bandwidth
curves of the other three memory technologies appear to be very different. The
reason is that write latencies of the three non-volatile memories are much
higher than read latencies. Another observation from
figure~\ref{fig:memory-bw}(b) is that the curves cross to each other at
different locations. Based on these two observations, we would like to design a
bandwidth-aware hybrid memory hierarchy, which always provides the high memory
bandwidth with the given capacity. For example, we probably would like to use
1) SRAM when the memory capacity is under 256MB, 2) MRAM or RRAM between
256MB and 1GB (if only bandwidth is considered), and 3) PCRAM when the memory
cacity is over 1GB.

@@

Based on the parameters of different cache configurations collected from our
modified version of CACTI~\cite{CACTI}, we evaluate both pure SRAM-based and
hybrid cache hierarchy designs. In this section, we show how bandwidth-aware
memory hierarchy design method help with improving the system performance.

>>>>

\documentclass[10pt,letterpaper,singlecolumn]{article}
\usepackage{setspace}
\doublespacing
\usepackage{latex8}
\usepackage{times}
\usepackage{cite}
\usepackage{subfigure}
\usepackage{psfig}
\usepackage[pdftex]{graphicx}
\usepackage{epsfig}
\usepackage{pstricks}
\usepackage{fullpage}
\usepackage{amsthm}
\usepackage{graphics}
\usepackage{comment}
\usepackage[small,compact]{titlesec}
\usepackage[small,it]{caption}
\usepackage{wrapfig}
\usepackage{enumerate}
\usepackage{multirow}

@@

First of all, we estimate the read and write bandwidths that can be provided by different memory technologies. Figure~\ref{fig:memory-bw} shows the results, with both x- and y-values in \emph{log} scale. The figure illustrates both the
provided read and write bandwidth as a function of memory capacity. Each of the memory technologies actually provide nearly the same read bandwidths, as is shown in figure~\ref{fig:memory-bw}(a). On the other hand, a straight forward observation from figure~\ref{fig:memory-bw}(b) is that the write bandwidth varies among different memory technologies. The shape of the SRAM write bandwidth curve is very similar to the read bandwidth curve. The write bandwidth curves of the other three memory technologies appear to be very different. The reason is that write latencies of the three non-volatile memories are much higher than read latencies. Another observation from figure~\ref{fig:memory-bw}(b) is that the curves cross to each other at different locations.

\begin{figure*}[htbp]
% The "!" means to maintain the aspect ratio.
\centering
\includegraphics[width=6in]{figures/RAM_energy}
\caption{Dynamic energy consumption with the provided bandwidths of different
  memory technologies. (a) Dynamic Energy with 20\% write.  (b) Dynamic Energy with 50\% write. }
\label{fig:memory-energy}
\includegraphics[width=6in]{figures/RAM_latency}
\caption{Latency of different memory technologies. (a) Read Latency. (b) Write Latency.}
\label{fig:memory-latency}
\end{figure*}

Being aware that the read and write latencies of NVM are asymmetric, we consider the read latency at first. In NVsim, we divide the entire cache read latency into the following components:
\begin{itemize}
\item (a)	H-tree input delay
\item (b)	Decoder + word-line delay
\item (c)	Bit-line delay
\item (d)	Sense Amplifier delay
\item (e)	Comparator Delay (for tag part only)
\item (f)	H-tree output delay
\end{itemize}
H-tree latency (a) and (f) are mainly determined by the RC delay of global wires, which is positive proportional to the area of memory macro. Sensing delay (d) is affected by the read noise margin of memory cell that is affected by off/on resistance ratio.

Based on these two observations, we would like to design a
bandwidth-aware hybrid memory hierarchy, which always provides the high memory
bandwidth with the given capacity. For example, we probably would like to use
1) SRAM when the memory capacity is under 256MB, 2) MRAM or RRAM between
256MB and 1GB (if only bandwidth is considered), and 3) PCRAM when the memory
cacity is over 1GB.

@@

 We will also use
Simics~\cite{Magnusson:2002:simics-orig-paper} as the simulator to evaluate our
design method. The benchmarks will be selected from bandwidth-bound
applications, which possibly include benchmarks from SPEC
CPU2006~\cite{SPEC:2006} and NPB~\cite{NPB}.\vspace{0.15in}
